Merging merges, more or less

### 2015-04-09 2015mergingmerges/section/01-Introduction.tex
add conclusion
rough figure of group by

% Most dataflow language optimisations tend to focus on fully static networks where the exact access pattern is known at compile time\cite{thies2002streamit}, but disallowing combinators like merge join, append and filter.
% At the other side of the spectrum, some dataflow languages such as Lucid\cite{stephens1997survey} focus on expressivity, forgoing any kind of static analyses for optimisations.

% Merge joins, appends, and filters are not typically necessary for the bulk kind of operations such as audio transforms, video compression and so on, that regular dataflow has found its applications in\cite{johnston2004advances}.

% For querying large data sets merge-like combinators are essential.
% The ``MapReduce'' framework has been touted as a solution to easy distribution of workloads, but has been found to be lacking in flexibility\cite{vrba2009kahn}, in favour of Kahn process networks.
% Kahn process networks alone are too flexible: many interesting properties we would like to assure, such as the absence of deadlocks, and ability to run in bounded memory, are undecidable.
% Regular dataflow languages correspond to a subset of Kahn process networks, restricted to the point of keeping these desired properties\cite{thies2009language}.

% In functional languages, fusion systems such as stream fusion\cite{coutts2007stream} and fold/build fusion\cite{jones2001playing} rely on the inliner to move producers into their consumers, after which rewrite rules can remove intermediate allocations.
% This short-cut fusion works well for vertical fusion when producers have only single consumers, but when a producer is used by multiple consumers it cannot be inlined without duplicating work, so fusion will not occur.
% Our earlier work on flow fusion\cite{lippmeier2013data} is able to fuse these multiple consumer cases, but only for a small set of combinators; neither @append@ nor @merge@ are handled.

% Our goal is to extend the regular dataflow languages enough to allow this subset of dynamic combinators, while keeping the desired properties, for the purpose of compilation and optimisation.

For example, if we have two sets of values over time, one useful operation is a correlation: to find the intersection of the values at given times, and compute some new value based on the two values:
\begin{code}
correlate :: [(Date,Value)] -> [(Date,Value)]
          -> [(Date,Value)]
correlate c1 c2
 = let m1 = merge c1 c2
       g1 = group_by compute m1
   in  g1
\end{code}

@@

\begin{code}
corr2     :: [(Date,Value)] -> [(Date,Value)]
          -> [(Date,Value)]
          ->([(Date,Value)],   [(Date,Value)])
corr2     c1 c2 c3
 = let m1 = merge c1 c2
       m2 = merge c2 c3
       g1 = group_by compute m1
       g2 = group_by compute m2
   in (g1, g2)
\end{code}

>>>>

For example, if we have two sets of values over time, one useful operation is a correlation: to find the intersection of the values at given times, and compute some new value based on the two values:
\begin{code}
correlate :: [(Date,Value)] -> [(Date,Value)]
          -> [(Date,Value)]
correlate c1 c2
 = let m1 = merge c1 c2
       g1 = groupby compute m1
   in  g1
\end{code}

@@

\begin{code}
corr2     :: [(Date,Value)] -> [(Date,Value)]
          -> [(Date,Value)]
          ->([(Date,Value)],   [(Date,Value)])
corr2     c1 c2 c3
 = let m1 = merge c1 c2
       m2 = merge c2 c3
       g1 = groupby compute m1
       g2 = groupby compute m2
   in (g1, g2)
\end{code}

### 2015-04-09 2015mergingmerges/section/02-Machines.tex
remove psis
rough figure of group by

\begin{tabbing}
MMMM \= MM \= MMMMMM \= M \= M \kill
$@Type@$ \> $=$ \> @Pull@     \>     \> $n$         \\
         \>     \> (Attempt to read from a named input) \\

         \> $~|$\> @Release@  \>     \> $n$         \\
         \>     \> (Values must be released before the input can be pulled again) \\

         \> $~|$\> @Close@    \>     \> $n$         \\
         \>     \> (A named input may still have data, but we will not read it) \\

@@

Many interesting combinators can be described using these machines.
In figures~\ref{fig:com:map}-\ref{fig:com:merge}, the machines for some important combinators are given.
To reduce clutter, these figures omit labels for @Unit@ transitions.

>>>>

\begin{tabbing}
MMM \= MM \= MMMMMM \= M \= M \kill
$@Type@$ \> $=$ \> @Pull@     \>     \> $n$         \\
         \>     \> (Attempt to read from a named input) \\

         \> $~|$\> @Release@  \>     \> $n$         \\
         \>     \> (Values must be released before the next pull) \\

         \> $~|$\> @Close@    \>     \> $n$         \\
         \>     \> (An input may still have data, but we will not read it) \\

@@

Many interesting combinators can be described using these machines.
In figures~\ref{fig:com:map}-\ref{fig:com:groupby}, the machines for some important combinators are given.
To reduce clutter, these figures omit labels for @Unit@ transitions.
